---
title: "Machine_learning"
author: "Martin Polacek"
date: "25 Jul 2015"
output: html_document
---
#INTRODUCTION
In this project we analyse the data from the webpage "http://groupware.les.inf.puc-rio.br/har", section "Weight Lifting Exercises Dataset" (for more info see also [1]). The dataset describes the quality of the exercise of 6 people. The goal of this project is to predict the manner in which they did the exercise.

#SETUP 
We download the required data and load libraries:
```{r, echo=T, cache=TRUE}
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv","./training.csv",method="curl")
```

```{r, echo=T, cache=TRUE}
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv","./testing.csv",method="curl")
```

```{r, echo=T, cache=TRUE}
training<-read.csv("training.csv")
```

```{r, echo=T, cache=TRUE}
test<-read.csv("testing.csv")
```

```{r, echo=F, cache=TRUE}
library(caret)
```

Install packages for the parallel computation and the dplyr package for easier data handling:

```{r, echo=F, cache=TRUE}
library(doParallel)
registerDoParallel(cores=2)
```

```{r, echo=F, cache=TRUE}
library(dplyr)
```

#ANALYSIS 

We clear the "training set". Brief look at the data reveals that there are many columns with NA values or columns used for a book keeping. We get rid of those first.

```{r, echo=T, cache=TRUE}
training <- subset(training, select = c("roll_dumbbell","pitch_dumbbell","yaw_dumbbell","roll_arm","pitch_arm","yaw_arm","total_accel_arm","roll_belt","pitch_belt","yaw_belt","total_accel_belt","accel_dumbbell_z","magnet_dumbbell_z","roll_forearm","pitch_forearm","yaw_forearm","accel_forearm_z","magnet_forearm_z","classe"))

```

We subdivide the "training" data into the two groups "train1", "train2" with the following percentages (respectivelly) "90%","10%". The "train1" set is used as a genuine training set. The "train2" determine the out of sample error, i.e. the probablility to misclassify the class.

```{r, echo=T, cache=TRUE}
sub1<-createDataPartition(y=training$classe,p=0.9,list=F)
sub2<-createDataPartition(y=training$classe,p=0.1,list=F)
train1<-training[sub1,]
train2<-training[sub2,]
```

##TRAINING

Since we are dealing with the classification problem, we can use either a simple "decision tree" or more sophisticated "random forest". We go for the "random forest".

We compute the "random forest" for the "train1" data set:

```{r, echo=T, cache=TRUE}
set.seed(835)
train_var <-train(classe~.,data=train1,method="rf",ntree=10)
```

the accuracy of "train_var" is arround "97%"

```{r, echo=T, cache=TRUE}
train_var
```

We save the trained model:

```{r, echo=T, cache=TRUE}
saveRDS(train_var, file="train_var.rds")
```

##TESTING on "train2"

The train_var has already done cross validations and the accuracy is "97%". We would like to test the results on the independent data set "train2". We can thus independently estimate the out of sample error.

```{r, echo=T, cache=TRUE}
res_sub2<-predict(train_var,train2)
```

```{r, echo=T, cache=TRUE}
err = function(val, pred) {
    sum(pred != val)/length(val)
}
out_of_sample_error = err(train2$classe,res_sub2)
```

The out of sample error is "out_of_sample_error = 0.0015"

##REFERENCES

[1] Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.
 

